The contribution of drifting snow to cloud properties and the atmospheric radiative budget over Antarctica

### 2021-10-14 submission/review_1/agujournaltemplate.tex
Review 1

%% To submit your paper:
\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\usepackage{float}
\usepackage[T1]{fontenc}
\linenumbers
%%%%%%%
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%

@@

\affiliation{1}{Department of Geosciences, University of Oslo, Oslo, Norway}
\affiliation{2}{Department of Geography, SPHERES Research Unit, University of Liège, Liège, Belgium}
\affiliation{3}{Univ. Grenoble Alpes, CNRS, Institut des Géosciences de l’Environnement, Grenoble, France}
\affiliation{4}{Univ. Grenoble Alpes, Université de Toulouse, Météo-France, CNRS, CNRM, Centre d’Études de la Neige, Grenoble, France}
%(repeat as many times as is necessary)

@@

\begin{abstract}
The Antarctic Ice Sheet experiences perpetual katabatic winds, transporting snow and moisture from the interior towards the periphery. However, the impacts of Antarctic moisture and drifting snow on cloud structure and surface energy fluxes have not been widely investigated. Here, we use a regional climate model with a newly-developed drifting snow scheme to show that accounting for drifting snow notably alters the spatial distribution, vertical structure and radiative effect of clouds over Antarctica. Overall, we find that accounting for drifting snow leads to a greater cloud cover providing an increase of +2.74\,Wm\textsuperscript{-2} in the surface radiative energy budget. Additionally, a comparison with 20 weather stations reveals a 2.17\,Wm\textsuperscript{-2} improvement in representing the radiative energy fluxes. Our results highlight the need to study the impact of missing drifting snow processes on the future evolution of clouds, the surface energy budget and the vertical atmospheric structure over Antarctica.
\end{abstract}

\section*{Plain Language Summary}
Antarctica is the continent with the strongest winds on Earth. These winds pick up a lot of snow on their way from the interior towards the ocean, forming drifting snow clouds. Drifting snow clouds can extend over 1000 km horizontally and multiple 100 m vertically. Like a normal cloud, they can reflect incoming sunlight like a mirror and trap heat like a blanket. However, most of our climate models don't yet incorporate these drifting snow clouds and therefore might be missing an important part of the Antarctic climate system. In this study we show that when we account for drifting snow clouds the Antarctic surface receives notably more thermal radiation. Additionally, we also show that we significantly improve our model when we include blowing snow by comparing our outputs to weather station observations over Antarctica. Therefore, we conclude that accurate sea level rise projections need to account for drifting snow.

@@

Clouds are known to notably affect the present and future climates of polar ice sheets \cite{Gilbert2020,Gorodetskaya2015,Lachlan2010,Hofer2017, Hofer2019,Hahn2019}. They have the ability to amend incoming and outgoing shortwave and longwave fluxes, depending on the cloud phase, height and particle size distribution, directly impacting the surface energy budget \cite{Gilbert2020,Tan2016,Tan2019}. Optically-thick drifting-snow clouds, while not accounted for in most global and regional climate models, can change the atmospheric radiation budget \cite{Letoumelin2020}, most notably because drifting-snow layers act as a cloud themselves, increasing the atmospheric longwave emissivity and decreasing the shortwave transparency of the atmosphere \cite{Yang2013, Yamanouchi1984, Letoumelin2020, Lawson2006}. Further, drifting-snow sublimation acts as a moisture source and a heat sink and therefore changes the temperature and humidity distribution in the near-surface atmosphere \cite{Amory2019}. Additionally, drifting-snow particles can also act as ice nucleating particles for cloud formation \cite{Geerts2015}, which impact the longevity, structure, cloud-phase distribution and precipitation formation within pre-existing clouds. While the near-surface air temperatures in the interior of Antarctica are often below -37\,\textdegree C, where homogenous cloud droplet freezing glaciates all clouds, above the boundary layer mixed-phase clouds can still exist in the Antarctic interior \cite{Lawson2014}, which are susceptible to changes in available ice nuclei. However, so far very little is known about how clouds are influenced by drifting-snow processes in climate models, and how accounting for drifting snow over the current climate influences key polar cloud-, and therefore climate processes.

Here, we use two regional climate model simulations spanning the period of 2000-2019, one with a dynamic representation of drifting snow and one without, to assess the impact of accounting for drifting snow on the representation of Antarctic clouds and surface radiative fluxes. We compare our two simulations during the 2000-2019 period to concurrently available satellite products of cloud cover and the ERA5 reanalysis, to show whether accounting for drifting snow only amends or also improves the comparison of modelled to observed Antarctic clouds. However, current satellite cloud cover products over Antarctica often do not report values in the lowermost 100 m  of the atmosphere where drifting snow is frequent and therefore neglect drifting snow and render model cloud cover verification difficult. Nevertheless, our results deliver a clear indication that accounting for drifting snow over polar ice sheets changes the 3D-structure of clouds and ultimately their contribution to the surface energy budget. Due to their similarity in radiative effects and also particle size \cite{Lawson2006}, we think that thick drifting-snow layers should be referred to as drifting-snow clouds and be included in satellite products used for model cloud cover evaluation. In conclusion, not accounting for drifting snow in future projections of the Antarctic climate and sea level rise contribution might significantly bias the drawn conclusions.

@@

In this study, we used the latest model version of MAR (v3.11), which includes a recently updated drifting-snow scheme described and evaluated in \citeA{amory2021}. Erosion of snow in the model occurs when the wind shear stress exerted at the surface exceeds a threshold value that depends only upon surface snow density ($\rho_{s}$) when $\rho_{s}$ < 450 kg/m\textsuperscript{3}.

@@

\subsection{CloudSat-CALIPSO cloud fraction}
For the comparison of the cloud cover simulated by MAR with satellite observations, we use the combined CloudSat spaceborne radar and CALIPSO spaceborne lidar cloud fraction dataset \cite{kay2009}. It is based on the R04 versions of the CloudSat standard products 2B-GEOPROF \cite{marchand2008} and 2B-GEOPROF-LIDAR \cite{mace2009} and provides the cloud fraction globally (82°S-82°N) on a 2°\,x\,2° horizontal grid with a 480\,m vertical resolution. The great advantage of using this active remote sensing dataset is its independence from the surface albedo over the bright Antarctic \cite{kay2016}. Here, we use the total mean cloud fraction between July 2006 and February 2011. Note however, that the widely-used CloudSat-CALIPSO cloud fraction product ignores cloud cover below 720 m, the part of the atmosphere where drifting-snow clouds are most frequently observed.

@@

Explicitly modelling drifting snow in MAR leads to a notable change in the atmospheric structure of the lowermost 100s of meters above ground (Fig.\ref{fig:Test} A-C). Over the flat interior of the Antarctic Ice Sheet, the first few 100\,m show a strong decrease in atmospheric temperature, with a mean 0-500\,m difference of -0.66 $\pm$ 0.40\,\textdegree C in elevations greater than 2000\,m above mean sea level (Fig.\ref{fig:Test} A). Conversely, over the lower grounded ice and the low-lying ice shelves surrounding the Antarctic Ice Sheet (<100\,m above sea level), this decrease in temperature in the drifting snow simulations is less notable. The mean 0-500\,m above surface difference lies at -0.23 $\pm$ 0.15\,\textdegree C. The contrasting picture between the flat interior and the steeper and lower margins of Antarctica is likely caused by a contrast in atmospheric turbulence: 1) Due to the low height gradients over the interior plateau and the corresponding stable boundary layer and less pronounced effect of turbulent mixing, the sublimational cooling is not mixed as efficiently as over the steeper margins. Therefore, we see a stronger boundary layer cooling in the interior when accounting for drifting snow sublimation, despite lower total erosion of snow by the wind than over steeper terrain. Sublimation cools the atmosphere because the change of water phase from solid to gaseous requires energy from the surrounding air to break up the bonds between the H$_2$O molecules, leading to a drop in temperature. 2) Due to adiabatic warming and strong turbulent mixing in areas where the gravitational pull accelerates the katabatic winds down steep terrain, the height of the boundary layer increases and the particles are entrained into higher elevations. Therefore, the sublimational cooling is less concentrated over the margins of Antarctica and the ice shelves, despite a greater sublimation potential due to higher temperatures and increased erosion fluxes over the steeper margins.

@@

In the boundary layer, accounting for drifting snow also increases cloud ocurence over the Antarctic continent (Fig.\ref{fig:Test}B). Our results show that the strongest increase in 2000-2019 average cloud cover over the interior plateau strongly overlap with the changes in temperature seen in Figure \ref{fig:Test}A. In elevations above 2000\,m above mean sea level the lowermost 500\,m of the atmosphere show an increase of +18.4 $\pm$ 11.8\% in cloud cover. Again, over lower elevations (<100 m) the signal is less pronounced, with an increase in cloud cover of +12.5 $\pm$ 8.4\%.

Generally, there are three overlapping mechanisms that can explain the greater cloud amount over Antarctica, when accounting for drifting snow. 1) Thick drifting-snow layers themselves act as a cloud, due to their ability to interact with incoming solar radiation (i.e. a cloud optical depth > 0) and their influence on the atmospheric longwave emissivity (i.e. they increase the atmospheric longwave emissivity $\epsilon$). 2) The sublimation of airborne snow particles leads to a cooling of the surrounding air, while increasing the specific humidity, both bringing the environment closer to saturation \cite{Amory2019}. 3) Drifting snow particles can act as additional nuclei on which water vapor can sublimate or help with ice growth through the Wegener-Bergeron-Findeisen process in mixed-phase clouds above the boundary layer. Ice crystal number concentration can furthermore potentially multiply through secondary ice processes \cite{Soti2020}. It is likely that in most cases these three processes can act simultaneously.

@@

 Conversely, over the drier and colder interior of Antarctica, we see virtually no changes in liquid water path despite a significant increase in cloud cover (Fig.\ref{fig:micro}A-C). However, our results suggest a widespread increase in cloud ice water path (Fig.\ref{fig:micro}D), with a mean increase over the grounded AIS of +5.9\,g/m\textsuperscript{-2} and even more over the ice shelves with an increase of +9.1\,g/m\textsuperscript{-2} in MAR with drifting snow. These changes in cloud ice water path correspond to a +10.3\% increase over the grounded ice and a 10.2\% increase over the ice shelves  %Additionally, new ice particles can also grow in MAR due to higher relative humidity of the advected air via the gas phase or via heterogeneous and homogeneous freezing \cite{Gallee1994}.
 Note however, that the MAR cloud microphysics scheme currently does not account for secondary ice production, where one single ice crystal can turn into multiple ice crystals via collision breakup, drop shattering and rime splintering \cite{Gallee1994, Storelvmo2015, Soti2020, Field2017}. However, especially rime splintering and drop shattering need liquid to be present and are most efficient in temperatures above what we observe over Antarctica \cite{Soti2020}. Therefore, we do not think that the missing drop shattering and rime splintering processes are a major source of uncertainty in our simulations, however, collision breakup in drifting-snow clouds could be an important missing multiplier of ice crystal number concentration in our simulations.

@@

To find out how adding active drifting snow to MAR changes the modelled cloud cover, we compare MAR and ERA5 to the Cloudsat-Calipso active satellite cloud cover product \cite{kay2009, marchand2008, mace2009} (Fig.\ref{fig:sat} A-C). Over the periods where Cloudsat-Calipso data is available (07/2006-02/2011), we find that MAR without active drifting snow overestimates cloud cover by 7.9 $\pm$ 9.2\,\% (Fig.\ref{fig:sat} A). The slight overestimation seems to be enhanced over East Antarctica. Furthermore, MAR with active drifting snow increases the overestimation of cloud cover to 25.4$\pm$ 12.4\,\% (Fig.\ref{fig:sat} B). Otherwise, MAR with drifting snow shows a spatially homogenous bias with little spatial variability. For a better understanding where MAR cloud cover biases rank compared to the widely used state-of-the-art reanalysis product ERA5 \cite{Hersbach2020}, we also compare ERA5 to the Cloudsat-Calipso cloud cover product. ERA5 shows a slightly larger overestimation of cloud cover (9.8 $\pm$ 14.5\,\%) than MAR without drifting snow, but 15.6\% less than MAR with drifting snow.

@@

Note however, that even though the global gridded CloudSat-CALIPSO cloud cover product here is one of the most advanced cloud products available for comparison with climate models, it does not include information about cloud cover below 720 m above the surface \cite{kay2009}. Therefore, because drifting-snow clouds are mostly less than 500 m in vertical extent \cite{Palm2018}, it is hard to assess with the currently available products whether accounting for drifting snow in MAR improves or degrades the performance with respect to cloud cover. Further, while there is only limited observational evidence for the size distribution of drifting snow particles, a case study over the South Pole station found that drifting snow particles are mostly between 30~{\textmu}m and 100~{\textmu}m in size \cite{Lawson2006}, a range also observed for typical cloud ice crystals. This similarity likely indicates that drifting snow clouds have similar optical and radiative properties to "conventional" clouds, and therefore information about drifting-snow clouds should be added to satellite cloud cover products over Antarctica.

@@

\begin{figure}[H]
	\includegraphics[scale=0.65]{SEB.png}
	\caption{\textbf{Difference in radiative components at the surface between MAR with and without drifting snow.} A) Difference in incoming shortwave radiation (SWD) at the surface in $Wm^{-2}$. Red color indicates a greater downwelling shortwave flux in MAR with active drifting snow parameterisation. B) Same A) but for the downwelling longwave flux at the surface. C) Same as A) and B), but for the difference in the net radiation at the surface ($R= SWD * (1 - \alpha) + LWD - LWU$).}
	\label{fig:SEB}
\end{figure}

@@

When looking at the net radiative effect of drifting snow (Fig.\ref{fig:SEB}C), we see that including drifting snow leads to a net radiative warming of +2.74\,Wm\textsuperscript{-2} over the grounded Antarctic Ice Sheet and +1.43\,Wm\textsuperscript{-2} over the ice shelves. Here, the radiative warming effect is mostly caused by an increase in LWD, most notably over the steep margins, and by a decrease in outgoing longwave radiation due to sublimation of drifting-snow particles cooling the near surface atmosphere. We find only limited evidence for a notable contribution of net shortwave radiation through changes in the surface albedo when accounting for drifting snow (not shown). Over the steeper terrain we see an increase in cloud cover, together with the strongest increase in cloud ice water path due to greater wind speeds and snow erosion, causing an enhanced atmospheric longwave emissivity (Fig.\ref{fig:micro}D).

@@

When comparing MAR to 20 in-situ weather station observations across the Antarctic Ice Sheet, the mean bias is notably reduced in our simulation with active drifting snow (Fig. \ref{fig:heat}, the mean bias for individual stations can be found in Supplementary Fig.~1). The reduction of the mean bias in absolute terms is greatest in the longwave part of the spectrum with -1.1\,Wm\textsuperscript{-2} in the downwelling longwave radiation (LWD) and -1.6\,Wm\textsuperscript{-2} in the outgoing longwave radiation (LWU, Fig. \ref{fig:heat} first row). Additionally, MAR with drifting snow has no notable impact on the outgoing shortwave radiation (SWU), where the mean bias is almost constant at +0.07\,Wm\textsuperscript{-2}, while it is slightly increased in the downwelling shortwave component (SWD) at +0.46\,Wm\textsuperscript{-2}. Overall, acounting for drifting snow in MAR over Antarctica leads to a 2.17\,Wm\textsuperscript{-2} better representation of the radiative fluxes when compared to observations (-1.6 - 1.1 +0.07 + 0.46 = -2.17\,Wm\textsuperscript{-2}). The greatest improvement in the mean bias is related to the two longwave components of the surface energy budget when explicitly modelling drifting snow over Antarctica.

@@

Our results also answer the question whether accounting for drifting snow leads to a net positive or negative radiative effect over Antarctica. We find that drifting snow leads to a net radiation increase at the surface of +2.74\,Wm\textsuperscript{-2} over the grounded parts of the Antarctic Ice Sheet, which could ultimately contribute to global sea level rise (Fig. \ref{fig:SEB}). Note however, that more local analysis suggests that a decrease in the latent heat flux through sublimation cooling might partly offset some of the radiative warming \cite{Letoumelin2020}.

@@

\subsubsection*{Code and data availability}

>>>>

%% To submit your paper:
\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\usepackage{float}
\usepackage{xcolor}
%\usepackage[T1]{fontenc}
\linenumbers
%%%%%%%
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%

@@

\affiliation{1}{Department of Geosciences, University of Oslo, Oslo, Norway}
\affiliation{2}{Laboratory of Climatology, SPHERES Research Unit, University of Liège, Liège, Belgium}
\affiliation{3}{Univ. Grenoble Alpes, CNRS, Institut des Géosciences de l’Environnement, Grenoble, France}
\affiliation{4}{Univ. Grenoble Alpes, Université de Toulouse, Météo-France, CNRS, CNRM, Centre d’Études de la Neige, Grenoble, France}
%(repeat as many times as is necessary)

@@

\begin{abstract}
The Antarctic Ice Sheet experiences perpetual katabatic winds, transporting snow and moisture from the interior towards the periphery. However, the impacts of Antarctic moisture and drifting snow on cloud structure and surface energy fluxes have not been widely investigated. Here, we use a regional climate model with a newly-developed drifting snow scheme to show that accounting for drifting snow notably alters the spatial distribution, vertical structure and radiative effect of clouds over Antarctica. Overall, we find that accounting for drifting snow leads to a greater cloud cover providing an increase of +2.74\,Wm\textsuperscript{-2} in the surface radiative energy budget. Additionally, a comparison with 20 weather stations reveals a 2.17\,Wm\textsuperscript{-2} improvement in representing the radiative energy fluxes. Our results highlight the need to study the impact of drifting snow processes on the future evolution of clouds, the surface energy budget and the vertical atmospheric structure over Antarctica.
\end{abstract}

\section*{Plain Language Summary}
Antarctica is the continent with the strongest winds on Earth. These winds pick up a lot of snow on their way from the interior towards the ocean, forming drifting snow clouds. Drifting snow clouds can extend over 1000 km horizontally and multiple 100 m vertically. Like a normal cloud, they can reflect incoming sunlight like a mirror and trap heat like a blanket. However, most of our climate models don't yet incorporate these drifting snow clouds and therefore might be missing an important part of the Antarctic climate system. In this study we show that when we account for drifting snow clouds the Antarctic surface receives notably more thermal radiation. Additionally, we also show that we significantly improve our model when we include drifting snow by comparing our outputs to weather station observations over Antarctica. Therefore, we conclude that accurate Antarctic climate projections need to account for drifting snow.

@@

Clouds are known to notably affect the present and future climates of polar ice sheets \cite{Gilbert2020,Gorodetskaya2015,Lachlan2010,Hofer2017, Hofer2019,Hahn2019}. They have the ability to amend incoming and outgoing shortwave and longwave fluxes, depending on the cloud phase, height and particle size distribution, directly impacting the surface energy budget \cite{Gilbert2020,Tan2016,Tan2019}. Optically-thick drifting-snow clouds, while not accounted for in most global and regional climate models, can change the atmospheric radiation budget \cite{Letoumelin2020}, most notably because drifting-snow layers act as a cloud themselves, increasing the atmospheric longwave emissivity and decreasing the shortwave transparency of the atmosphere \cite{Yang2013, Yamanouchi1984, Letoumelin2020, Lawson2006}. Further, drifting-snow sublimation acts as a moisture source and a heat sink and therefore changes the temperature and humidity distribution in the near-surface atmosphere \cite{Amory2019}. Additionally, drifting-snow particles can also act as ice nucleating particles for cloud formation \cite{Geerts2015}, which impact the longevity, structure, cloud-phase distribution and precipitation formation within pre-existing clouds. While the near-surface air temperatures in the interior of Antarctica are often below -37\,\textdegree C, where homogenous cloud droplet freezing glaciates all clouds, mixed-phase clouds can still exist above the boundary layer in the Antarctic interior \cite{Lawson2014}, which are susceptible to changes in available ice nuclei. However, so far very little is known about how clouds are influenced by drifting-snow processes in climate models, and how accounting for drifting snow over the current climate influences key polar cloud-, and therefore climate processes.

Here, we use two regional climate model simulations spanning the period of 2000-2019, one with a dynamic representation of drifting snow and one without, to assess the impact of accounting for drifting snow on the representation of Antarctic clouds and surface radiative fluxes. We compare our two simulations during the 2000-2019 period to concurrently available satellite products of cloud cover and the ERA5 reanalysis, to show whether accounting for drifting snow only amends or also improves the comparison of modelled to observed Antarctic clouds. Our results deliver a clear indication that accounting for drifting snow over polar ice sheets changes the 3D-structure of clouds and ultimately their contribution to the surface energy budget. Due to their similarity in radiative effects and also particle size \cite{Lawson2006}, we think that thick drifting-snow layers should be referred to as drifting-snow clouds and be included in satellite products used for model cloud cover evaluation. In conclusion, not accounting for drifting snow in future projections of the Antarctic climate might notably bias the drawn conclusions.

@@

In this study, we used the latest model version of MAR (v3.11), which includes a recently updated drifting-snow scheme fully described and evaluated in \citeA{amory2021}. Erosion of snow in the model occurs when the wind shear stress exerted at the surface exceeds a threshold value that depends only upon surface snow density ($\rho_{s}$) when $\rho_{s} <$  450\,kg/m\textsuperscript{3}.

@@

\subsection{CloudSat-CALIPSO cloud fraction}
For the comparison of the cloud cover simulated by MAR with satellite observations, we use the combined CloudSat spaceborne radar and CALIPSO spaceborne lidar cloud fraction dataset \cite{kay2009}. It is based on the R04 versions of the CloudSat standard products 2B-GEOPROF \cite{marchand2008} and 2B-GEOPROF-LIDAR \cite{mace2009} and provides the cloud fraction globally (82°S-82°N) on a 2°\,x\,2° horizontal grid with a 480\,m vertical resolution. The great advantage of using this active remote sensing dataset is its independence from the surface albedo over the bright Antarctic \cite{kay2016}. Here, we use the total mean cloud fraction between July 2006 and February 2011.

CloudSat/CALIOP data was checked for cloud detection on a profile-by-profile basis. A positive cloud ID (meaning: cloud in this profile) requires a cloud thickness of 960\,m (480\,m for low clouds below 2.75\,km). CloudSat data below 720m a.s.l. are excluded due to surface clutter. Each individual profile is flagged this way as cloud/no-cloud, and the total cloud fraction is calculated as the number of cloudy profiles divided by the total number of profiles within the 2°x2° grid cell.

Note here, that it ignores cloud cover below 720 m, the part of the atmosphere where drifting-snow clouds are most frequently observed.

@@

Explicitly modelling drifting snow in MAR leads to a notable change in the atmospheric structure of the lowermost 100s of meters above ground (Fig.\ref{fig:Test} A-C). Over the flat interior of the Antarctic Ice Sheet, the first few 100\,m show a strong decrease in atmospheric temperature, with a mean 0-500\,m difference of -0.66 $\pm$ 0.40\,\textdegree C in elevations greater than 2000\,m above mean sea level (Fig.\ref{fig:Test} A, note: throughout the manuscript uncertainties are given as the mean spatial variability as $\pm$ 1 spatial standard deviation). Conversely, over the lower grounded ice and the low-lying ice shelves surrounding the Antarctic Ice Sheet ($<$100\,m above sea level), this decrease in temperature in the drifting snow simulations is less notable. The mean 0-500\,m above surface difference lies at -0.23 $\pm$ 0.15\,\textdegree C. The contrasting picture between the flat interior and the steeper and lower margins of Antarctica is likely caused by a contrast in atmospheric turbulence: 1) Due to the shallow surface slopes over the interior plateau and the corresponding stable boundary layer and less pronounced effect of turbulent mixing, the sublimational cooling is not mixed as efficiently as over the steeper margins. Therefore, we see a stronger boundary layer cooling in the interior when accounting for drifting snow sublimation, despite lower total erosion of snow by the wind than over steeper terrain. Sublimation cools the atmosphere because the change of water phase from solid to gaseous requires energy from the surrounding air to break up the bonds between the H$_2$O molecules, leading to a drop in temperature. 2) Due to adiabatic warming and strong turbulent mixing in areas where the gravitational pull accelerates the katabatic winds down steep terrain, the height of the boundary layer increases and the particles are entrained into higher elevations. Therefore, the sublimational cooling is less concentrated over the margins of Antarctica and the ice shelves, despite a greater sublimation potential due to higher temperatures and increased erosion fluxes over the steeper margins.

@@

In the boundary layer, accounting for drifting snow also increases cloud ocurence over the Antarctic continent (Fig.\ref{fig:Test}B). Our results show that the strongest increase in 2000-2019 average cloud cover over the interior plateau strongly overlap with the changes in temperature seen in Figure \ref{fig:Test}A. In elevations above 2000\,m above mean sea level the lowermost 500\,m of the atmosphere show an increase of +18.4 $\pm$ 11.8\% in cloud cover. Again, over lower elevations ($<$100 m) the signal is less pronounced, with an increase in cloud cover of +12.5 $\pm$ 8.4\%.

Generally, there are three overlapping mechanisms that can explain the greater cloud amount over Antarctica, when accounting for drifting snow. 1) Thick drifting-snow layers themselves act as a cloud, due to their ability to interact with incoming solar radiation (i.e. a cloud optical depth $>$ 0) and their influence on the atmospheric longwave emissivity (i.e. they increase the atmospheric longwave emissivity $\epsilon$). 2) The sublimation of airborne snow particles leads to a cooling of the surrounding air, while increasing the specific humidity, both bringing the environment closer to saturation \cite{Amory2019}. 3) Drifting snow particles can act as additional nuclei on which water vapor can sublimate or help with ice growth through the Wegener-Bergeron-Findeisen process in mixed-phase clouds above the boundary layer. Ice crystal number concentration can furthermore potentially multiply through secondary ice processes \cite{Soti2020}. It is likely that in most cases these three processes can act simultaneously.

@@

 Conversely, over the drier and colder interior of Antarctica, we see virtually no changes in liquid water path despite a significant increase in cloud cover (Fig.\ref{fig:micro}A-C). However, our results suggest a widespread increase in cloud ice water path (Fig.\ref{fig:micro}D), with a mean increase over the grounded AIS of +5.9\,g/m\textsuperscript{-2} and even more over the ice shelves with an increase of +9.1\,g/m\textsuperscript{-2} in MAR with drifting snow. These changes in cloud ice water path correspond to a +10.3\% increase over the grounded ice and a 10.2\% increase over the ice shelves.  %Additionally, new ice particles can also grow in MAR due to higher relative humidity of the advected air via the gas phase or via heterogeneous and homogeneous freezing \cite{Gallee1994}.
 Note however, that the MAR cloud microphysics scheme currently does not account for secondary ice production, where one single ice crystal can turn into multiple ice crystals via collision breakup, drop shattering and rime splintering \cite{Gallee1994, Storelvmo2015, Soti2020, Field2017}. However, especially rime splintering and drop shattering need liquid to be present and are most efficient in temperatures above what we observe over Antarctica \cite{Soti2020}. Therefore, we do not think that the missing drop shattering and rime splintering processes are a major source of uncertainty in our simulations, however, collision breakup in drifting-snow clouds could be an important missing multiplier of ice crystal number concentration in our simulations.

@@

We compare MAR and ERA5 to the Cloudsat-Calipso active satellite cloud cover product \cite{kay2009, marchand2008, mace2009} (Fig.\ref{fig:sat} A-C). Over the periods where Cloudsat-Calipso data is available (07/2006-02/2011), we find that MAR without active drifting snow overestimates cloud cover by 7.9 $\pm$ 9.2\,\% (Fig.\ref{fig:sat} A). The slight overestimation seems to be enhanced over East Antarctica. Furthermore, MAR with active drifting snow increases the overestimation of cloud cover to 25.4$\pm$ 12.4\,\% (Fig.\ref{fig:sat} B). Otherwise, MAR with drifting snow shows a spatially homogenous bias with little spatial variability. For a better understanding where MAR cloud cover biases rank compared to the widely used state-of-the-art reanalysis product ERA5 \cite{Hersbach2020}, we also compare ERA5 to the Cloudsat-Calipso cloud cover product. ERA5 shows a slightly larger overestimation of cloud cover (9.8 $\pm$ 14.5\,\%) than MAR without drifting snow, but 15.6\% less than MAR with drifting snow.

@@

Note however, that even though the global gridded CloudSat-CALIPSO cloud cover product here is one of the most advanced cloud products available for comparison with climate models, it does not include information about cloud cover below 720 m above the surface \cite{kay2009}. Therefore, because drifting-snow clouds are mostly less than 500 m in vertical extent \cite{Palm2018}, it is hard to assess with the currently available products whether accounting for drifting snow in MAR improves or degrades the performance with respect to cloud cover. Further, below 2.75 km Cloudsat-CALIPSO data requires a minimum cloud thickness of 480 m in vertical extent, notably limiting the usefulness of active satellite data for comparison with regional climate models that include drifting snow. Conversely, biases in cloud cover between satellite observations and our regional climate model could also be caused by different definitions of what constitutes a cloud. However, we conclude that even if we would include a satellite simulator in our model (such as COSP), we would not be able to compare our model output to observations in a meaningful way, because data below 720 m is excluded in the observations due to surface clutter, the height in which drifting snow clouds most frequently occur.

Additionally, while there is only limited observational evidence for the size distribution of drifting snow particles, a case study over the South Pole station found that drifting snow particles are mostly between 30~{\textmu}m and 100~{\textmu}m in size \cite{Lawson2006}, a range also observed for typical cloud ice crystals. This similarity likely indicates that drifting snow clouds have similar optical and radiative properties to "conventional" clouds, and therefore information about drifting-snow clouds should be added to satellite cloud cover products over Antarctica.

@@

\begin{figure}[H]
	\includegraphics[scale=0.65]{SEB_with_stations.png}
	\caption{\textbf{Difference in radiative components at the surface and snow particle ratio between MAR with and without drifting snow.} A) Difference in incoming shortwave radiation (SWD) at the surface in $Wm^{-2}$. Red color indicates a greater downwelling shortwave flux in MAR with active drifting snow parameterisation. B) Same A) but for the downwelling longwave flux at the surface. C) Same as A) and B), but for the difference in the net radiation at the surface ($R= SWD * (1 - \alpha) + LWD - LWU$). D) Same as above but for the difference in snow particle content (g/kg), a measure of airborne drifting snow particles. Dots show the locations of the weather stations in our statistical comparison in Fig.\ref{fig:heat}}
	\label{fig:SEB}
\end{figure}

@@

When looking at the net radiative effect of drifting snow (Fig.\ref{fig:SEB}C), we see that including drifting snow leads to a net radiative warming of +2.74\,Wm\textsuperscript{-2} over the grounded Antarctic Ice Sheet and +1.43\,Wm\textsuperscript{-2} over the ice shelves. Here, the radiative warming effect is mostly caused by an increase in LWD, most notably over the steep margins, and by a decrease in outgoing longwave radiation due to sublimation of drifting-snow particles cooling the near surface atmosphere. When looking at the climatological difference in airborne snow particles caused by drifting snow (Fig.\ref{fig:SEB}D) we see that the snow particles ratio is mostly enhanced over the steeper surface slopes of Antarctica, where the gravitational pull accelerates the katabatic winds. These constitute also the areas where the longwave warming is most enhanced in our simulation with drifting snow.

Our results further highlight the efficiency at which drifting snow enhances the atmospheric longwave emissivity. Overall, downwelling longwave radiation at the surface is a combination of atmospheric temperature and emissivity ($LWD=\epsilon\cdot T^{4}$). The fact that we see a notable increase in longwave radiation at the surface despite an atmospheric cooling strengthens the conclusion that drifting snow is a notable - and often neglected - component of the Antarctic radiation budget.

We find only limited evidence for a notable contribution of net shortwave radiation through changes in the surface albedo when accounting for drifting snow (not shown). Over the steeper terrain we see an increase in cloud cover, together with the strongest increase in cloud ice water path due to greater wind speeds and snow erosion, causing an enhanced atmospheric longwave emissivity (Fig.\ref{fig:micro}D).

@@

When comparing MAR to 20 in-situ weather station observations across the Antarctic Ice Sheet, the mean bias is notably reduced in our simulation with active drifting snow (Fig. \ref{fig:heat}, the mean bias for individual stations can be found in Supplementary Fig.~1, the location of the stations in Fig.\ref{fig:SEB}D and Fig.S3). The reduction of the mean bias in absolute terms is greatest in the longwave part of the spectrum with -1.1\,Wm\textsuperscript{-2} in the downwelling longwave radiation (LWD) and -1.6\,Wm\textsuperscript{-2} in the outgoing longwave radiation (LWU, Fig. \ref{fig:heat} first row). Additionally, MAR with drifting snow has no notable impact on the outgoing shortwave radiation (SWU), where the mean bias is almost constant at +0.07\,Wm\textsuperscript{-2}, while it is slightly increased in the downwelling shortwave component (SWD) at +0.46\,Wm\textsuperscript{-2}. Overall, acounting for drifting snow in MAR over Antarctica leads to a 2.17\,Wm\textsuperscript{-2} better representation of the radiative fluxes when compared to observations (-1.6 - 1.1 +0.07 + 0.46 = -2.17\,Wm\textsuperscript{-2}). The greatest improvement in the mean bias is related to the two longwave components of the surface energy budget when explicitly modelling drifting snow over Antarctica.

We also compared our MAR model results to observations only during drifting snow days at the location of a given in-situ weather station (Fig. S2). We find that during drifting snow days that the reduction in the longwave biases is even more pronounced, leading to a three times higher LWD bias reduction of -3.3\,Wm\textsuperscript{-2}, equivalent to a 50\% reduction in the mean bias. Furthermore, using the same MAR model setup and observations it has been shown that during drifting snow events differences in LWD can reach up to 60\,Wm\textsuperscript{-2}, far outside the uncertainty of in-situ observations \cite{Letoumelin2020}.

@@

Our results also answer the question whether accounting for drifting snow leads to a net positive or negative radiative effect over Antarctica. We find that drifting snow leads to a net radiation increase at the surface of +2.74\,Wm\textsuperscript{-2} over the grounded parts of the Antarctic Ice Sheet, which could ultimately contribute to global sea level rise (Fig. \ref{fig:SEB}). Note however, that a regional analysis of MAR in coastal Adelie Land suggests that sublimation cooling might partly offset some of the radiative warming at the surface \cite{Letoumelin2020}.

@@

\subsubsection*{Open Research}

%The monthly means used in this study, from 1980-2100 of all three MAR RCP 8.5 simulations are available via ftp://ftp.climato.be/fettweis/MARv3.9/Greenland/. In case daily outputs are required, these can be requested from Xavier Fettweis (xavier.fettweis@uliege.be) and Stefan Hofer (s.hofer@bristol.ac.uk).
%
\textbf{Data and code owned by the authors:} all the code used for the analysis in this study is available on https://github.com/stefan-hofer/Antarctica\_clouds, but will be made available via zenodo upon acceptance of the manuscript. All the 2000-2019 averages from our MAR simulations with and without drifting snow are available via \hfill \break  https://zenodo.org/record/5037197.
